### 2021-11-30 R Code/Randomise_Papers.Rmd.Rmd
initial commit

---
title: Randomly select papers for review
output:
  html_document: default
  pdf_document: default
---

***

## Summary
This document contains code for randomising paper ranks for the review on Dynamic Occupancy Models. For each of four search parameters, papers are stratified into 5 year time periods and randomly ranked. The top 25% of papers in each ranking category will be reviewed; papers which do not qualify for review will be replaced with the next highest rank.

***

```{r setup, echo = FALSE, message = FALSE}
library("knitr")
library("tidyverse")
library("readxl")

opts_knit$set(root.dir = "/Users/kelleherj/Documents/Research/QAEco/ODM_Review/")
```

```{r Read-in excel sheets, echo = FALSE, message = FALSE, warning = FALSE}
# From the excel review spreadsheet, read in only the identifying characteristics 
identifierSheet <- readxl::read_xlsx("Review Spreadsheet.xlsx", 
                                     sheet = "Identifying", skip = 1) %>%
  # Drop rows prior to start of data
  slice(4:nrow(.)) %>%
  # Only need google search query and the ID reference
  select(Review_ID, Query = `Google Scholar search terms`) 
# Repeat with the study metadata
studyDataSheet <- readxl::read_xlsx("Review Spreadsheet.xlsx", 
                                    sheet = "Study data", skip = 1) %>%
  slice(4:nrow(.)) %>%
  # Only need year data and ID reference
  select(Review_ID, Year) 
```

```{r Combine and merge, echo = FALSE, message = FALSE}
# Merge paper queries and metadata
paperList <- full_join(identifierSheet, studyDataSheet, 
                       by = "Review_ID") %>% 
  # Drop empty rows, excluded, and incidental papers
  filter(!is.na(Query) & Query != "INCIDENTAL" & Year != "EXCLUDE") %>%
  # Break into time periods 
  mutate(YearPeriod = case_when(Year %in% as.character(2000:2005) ~ "2000:2005",
                                Year %in% as.character(2005:2010) ~ "2005:2010",
                                Year %in% as.character(2010:2015) ~ "2010:2015",
                                Year %in% as.character(2015:2021) ~ "2015:2021"))
```

```{r Randomly rank papers within strata}
rankedPapers <- paperList %>%
  # Add random ranks for all papers
  mutate(randRank = sample(x = 1:nrow(.), size = nrow(.))) %>%
  # Re-rank within strata
  group_by(Query, YearPeriod) %>%
  arrange(Query, YearPeriod, randRank) %>%
  mutate(groupRank = row_number())

```