### 2022-07-29 Analysis/Outputs/confirmationStats.Rmd
restructure files

---
title: key stats for confirmation
output:
  html_document: default
  pdf_document: default
---

***

```{r setup, echo = FALSE, message = FALSE}
library("knitr")
library("tidyverse") 
library("readxl")
library("sf")
```

```{r readin}
reviewTable <- read_xlsx("Data/ConfirmationTable.xlsx")

nonSPOM <- reviewTable %>%
  filter(SearchTerm %in% c("Dynamic occupancy model", "Multi-season occupancy model"))

```

```{r locations}
reviewTable %>%
  group_by(Biorealm) %>%
  summarize(TotalPapers = n()) %>%
  arrange(-TotalPapers)

reviewTable %>%
  group_by(Taxa) %>%
  summarize(TotalPapers = n()) %>%
  arrange(-TotalPapers)

reviewTable %>%
  group_by(IUCN) %>%
  summarize(TotalPapers = n()) %>%
  arrange(-TotalPapers)

reviewTable %>%
  group_by(AuthorCountry) %>%
  summarize(TotalPapers = n()) %>%
  arrange(-TotalPapers)

reviewTable %>%
  separate_rows(ObjectiveCategories, sep = ", ") %>%
  group_by(ObjectiveCategories) %>%
  summarise(count = n()) %>% 
  mutate(proportion = count/58) %>%
  arrange(-proportion) %>%
  filter(ObjectiveCategories != 'invasives') %>%

  ggplot() +
  geom_col(aes(x = ObjectiveCategories, y = proportion), fill = 'dodgerblue4') +
  ylim(c(0,1)) +
  scale_x_discrete(limits = c('conservation', 'management', 'methods', 'research'),
                    labels = c('Conservation', 'Management', 'Methods', 'Research')) +
  labs(x = "", y = "Proportion of papers") +
  theme_minimal() +
  theme(axis.text = element_text(size = 15),axis.title = element_text(size = 15))

```

```{r split covariate column}
nonSPOM %>%
  separate_rows(CovTypes, sep = ", ") %>%
  group_by(CovTypes) %>%
  summarise(count = n()) %>% 
  mutate(proportion = count/38) %>%
  filter(proportion > 0.125) %>%
  arrange(proportion) %>%

  ggplot(aes(x = reorder(CovTypes, -proportion),
             y = proportion)) +
  geom_col(fill = "dodgerblue4") +
  ylim(c(0,1)) + 
  labs(title = "Covariates usage in non-SPOM models",
       x = "Covariate class",
       y = "Proportion of papers") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 15),
        axis.title = element_text(size = 15)) +
  scale_x_discrete(limits = c('land_cover', 'habitat', 'biotic_interaction',
                              'hydro', 'land_use', 'anthropogenic'),
                   labels = c('Land Cover', 'Habitat', 'Biotic Interactions',
                              'Hydrologic', 'Land Use', 'Anthropogenic'))
```

```{r prop infer}
reviewTable %>% 
  count(InfSpatial) %>%
  mutate(Prop = n/58)

reviewTable %>% 
  count(InfFuture)%>%
  mutate(Prop = n/58)

reviewTable %>% 
  count(InfPast)%>%
  mutate(Prop = n/58)

reviewTable %>% 
  count(Maps)%>%
  mutate(Prop = n/58)

reviewTable %>%
  count(Taxa) %>%
  arrange(-n)%>%
  mutate(Prop = n/58)

reviewTable %>%
  count(ModelSel) %>%
  mutate(Prop = n/58)

reviewTable %>%
  count(ModelEval) %>%
  mutate(Prop = n/58)

t<-reviewTable %>%
  separate_rows(CovTypes, sep = ", ") %>%
  filter(!CovTypes %in% c("connectivity", "habitat", "survey", "none"))

reviewTable %>%
  separate_rows(CovTypes, sep = ", ") %>%
  group_by(CovTypes) %>%
  summarise(count = n()) %>% 
  mutate(proportion = count/58) %>%
  arrange(-proportion)  %>%
  ggplot(aes(x = reorder(CovTypes, -proportion),
             y = proportion)) +
  geom_col(fill = "dodgerblue4") +
  ylim(c(0,0.5)) + 
  labs(title = "Covariates usage in non-SPOM models",
       x = "Covariate class",
       y = "Proportion of papers") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 15),
        axis.title = element_text(size = 15)) +
  scale_x_discrete(limits = c('land_cover', 'biotic_interaction',
                              'hydro', 'land_use', 'anthropogenic'),
                   labels = c('Land Cover', 'Biotic Interactions',
                              'Hydrologic', 'Land Use', 'Anthropogenic'))

```

```{r maps}
ecoregions <- st_read("Data/Ecoregions2017/Ecoregions2017.shp") %>%
  st_make_valid() %>%
  group_by(REALM) %>%
  summarise() %>%
  st_simplify(dTolerance = 10000)

ecoregions <- ecoregions %>%
  st_simplify(dTolerance = 20000)

reviewTable %>%
  group_by(Biorealm) %>%
  summarize(TotalPapers = n()) %>%
  arrange(-TotalPapers)

ggplot(data = ecoregions) +
  geom_sf(aes(fill = REALM))
```

```{r}
data.frame(selMethod = c("A priori", "Stepwise", "Bayesian", "Predefined alternatives"),
           Proportion = c(.44, .24, .22, .11)) %>%
  ggplot(aes(x = reorder(selMethod, -Proportion),
             y = Proportion)) +
  geom_col(fill = "dodgerblue4") +
  ylim(c(0,1)) + 
  labs(x = "Model Selection Method",
       y = "Proportion of papers") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 15),
        axis.title = element_text(size = 15))

data.frame(infer = c("Limited to study\nduration and area  ", "Spatial", "Future"),
           Proportion = c(.62, .12, .29)) %>%
  ggplot(aes(x = reorder(infer, -Proportion),
             y = Proportion)) +
  geom_col(fill = "dodgerblue4") +
  ylim(c(0,1)) + 
  labs(x = "Type of inference",
       y = "Proportion of papers") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 15),
        axis.title = element_text(size = 15))
```