### 2022-08-30 Writing/Randomisation/Randomise_Papers.Rmd
need to get some first-round results first to determine new methods

---
title: Randomly select papers for review
output:
  html_document: default
  pdf_document: default
---

***

## Summary
This document contains code for randomising paper ranks for the review on Dynamic Occupancy Models. For each of four search parameters, papers are stratified into 5 year time periods and randomly ranked. The top 25% or top 5 papers in each ranking category, whichever is larger, will be evaluated for inclusion; papers which do not qualify for review will be replaced with the next highest rank.

***

```{r setup, echo = FALSE, message = FALSE}
library("knitr")
library("tidyverse") 
library("readxl")

# Need seed set for reproducibility
set.seed(5378)
```

Read in the excel sheets and process out the header
```{r Read-in excel sheets, echo = FALSE, message = FALSE, warning = FALSE}
# From the excel review spreadsheet, read in only the identifying characteristics 
identifierSheet <- readxl::read_xlsx("Data/Review Spreadsheet.xlsx", 
                                     sheet = "Metadata", skip = 1) %>%
  # Drop rows prior to start of data
  slice(4:nrow(.)) %>%
  # Only need google search query and the ID reference
  select(Review_ID, Query = `Search terms`) 
# Repeat with the study metadata
studyDataSheet <- readxl::read_xlsx("Data/Review Spreadsheet.xlsx", 
                                    sheet = "Study data", skip = 1) %>%
  # Cut out non-data rows
  slice(4:nrow(.)) %>%
  # Only need year data and ID reference
  select(Review_ID, Year) 
```

Combine the two sheets, the former containing metadata and the latter containing information on the papers
```{r Combine and merge, echo = FALSE, message = FALSE}
# Merge paper queries and metadata
paperList <- full_join(identifierSheet, studyDataSheet, 
                       by = "Review_ID")

# Exclude some non-query papers and add year periods to the rest
includedPapers <- paperList %>%
  # Drop empty rows, excluded, and incidental papers
  filter(!is.na(Query) & Query != "INCIDENTAL" & Year != "EXCLUDE") %>%
  # Break into time periods 
  mutate(YearPeriod = case_when(Year %in% as.character(2000:2005) ~ "2000:2005",
                                Year %in% as.character(2005:2010) ~ "2005:2010",
                                Year %in% as.character(2010:2015) ~ "2010:2015",
                                Year %in% as.character(2015:2021) ~ "2015:2021"))
```

Assign each paper a random overll rank, then order the papers within their strata by the overall rank
```{r Randomly rank papers within strata}
# Randomly rank papers, then rank within the strata
rankedPapers <- includedPapers %>%
  # Add random ranks for all papers
  mutate(randRank = sample(x = 1:nrow(.), size = nrow(.))) %>%
  # Re-rank within strata
  group_by(Query, YearPeriod) %>%
  arrange(Query, YearPeriod, randRank) %>%
  mutate(GroupRank = row_number()) 
```

Highlight which papers are ranked highly enough to qualify for evaluation 
```{r Highlight papers to read}
# Identify which papers qualify to be read
highlightedPapers <- rankedPapers %>%
  group_by(Query, YearPeriod) %>%
  mutate(PercentRank =  GroupRank/max(GroupRank)) %>%
  mutate(ToRead = case_when(
    # If there are fewer than 5, read them all
    max(GroupRank) <= 5 ~ "YES",
    # If there are 20 or fewer, read the top 5
    max(GroupRank) <= 20 & GroupRank <= 5 ~ "YES",
    max(GroupRank) <= 20 & GroupRank >= 5 ~ "NO",
    # If there are 21 or more, read the top 25%
    max(GroupRank) >= 21 & PercentRank <= 0.25  ~ "YES",
    max(GroupRank) >= 21 & PercentRank > 0.25 ~ "NO")) %>%
  ungroup() %>%
  unite(col = Group, c(Query, YearPeriod)) %>%
  # Get Review_ID, GroupRank, and ToRead
  select(Review_ID, Group, OverallRank = randRank, GroupRank, ToRead)
```

Add previously unranked papers, remove extraneous columns, and save both this workspace and the rankings to disk.
```{r Reformat and save}
# Get the unranked papers to add to the list
unrankedPapers <- paperList %>% 
  # Drop empty rows, excluded, and incidental papers
  filter(is.na(Query) | Query == "INCIDENTAL" | Year == "EXCLUDE") %>%
  mutate(ToRead = "NO", Group = "Unranked", GroupRank = "NA", OverallRank = "NA") %>%
  select(Review_ID, Group, OverallRank, GroupRank, ToRead)

# Combine ranked and unranked papers
outputList <- rbind(highlightedPapers, unrankedPapers) %>%
  mutate(Review_ID = as.numeric(Review_ID)) %>%
  arrange(Review_ID)

# Write to disk
write_csv(outputList, file = "RankedPapers.csv")
# Save environment as well for reproducibility
save.image("Randomise_Papers.RData")
```