### 2023-01-26 Analysis/Code/2_InitialAnalysis.Rmd
first few basic figs

---
title: "DOM Review: Initial analysis"
output: html_document
---

```{r Setup, include = FALSE}
library(tidyverse)
```

```{r Readin, echo = FALSE}
reviewResults <- read_rds("Analysis/Outputs/ReviewResults.rds")
```

```{r Study, echo = FALSE}

```

```{r Objectives, echo = FALSE}

reviewResults$Objectives %>%
  pivot_longer(c("Obj_Methods","Obj_Projections","Obj_Spatial","Obj_Trends","Obj_Hypothesis"),
               "Objective") %>%
  mutate(value = as.numeric(value)) %>%
  group_by(Objective) %>%
  summarise(n=sum(value)) %>%

  ggplot(aes(x = Objective, y = n/54)) +
  geom_col() +
  ylim(0,1)
```

```{r Taxa, echo = FALSE}

```

```{r Survey, echo = FALSE}
hist(as.numeric(reviewResults$Survey$`Study duration`))

reviewResults$Survey %>%
  group_by(`Capture method`) %>%
  summarize(n = n()) %>%

  ggplot(aes(x = reorder(`Capture method`, -n), y = n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle=90, vjust = 0.5,hjust=1))

```

```{r Covariates, echo = FALSE}
nrow(reviewResults$Covariates)

reviewResults$Covariates %>%
  filter(!is.na(Category)) %>%
  group_by(Category, Scale) %>%
  summarize(Count = n()) %>%
  arrange(Count) %>%

  ggplot(aes(x = reorder(Category, -Count), y = Count, fill = Scale)) +
  geom_col() + 
  scale_fill_discrete(limits = c("a","p","d","g","u", NA), 
                    labels = c("area","point","distance","global", "Unclear", "NA")) +
  theme(axis.text.x = element_text(angle=90, vjust = 0.5,hjust=1))

reviewResults$Covariates %>%
  filter(!is.na(Category)) %>%
  group_by(Category, Dynanimity) %>%
  summarize(Count = n()) %>%
  arrange(Count) %>%

  ggplot(aes(x = reorder(Category, -Count), y = Count, fill = Dynanimity)) +
  geom_col() + 
  scale_fill_discrete(limits = c("d","s","u"), 
                    labels = c("Dynamic", "Static", "Unclear")) +
  theme(axis.text.x = element_text(angle=90, vjust = 0.5,hjust=1))

```

```{r Modelling, echo = FALSE}

```