### 2023-02-03 Analysis/Code/2_InitialAnalysis.Rmd
added some obj plots

---
title: "DOM Review: Initial analysis"
output: html_document
---

```{r Setup, include = FALSE}
library(tidyverse)
library(sf)
library(knitr)
```

## Summary
This document includes some preliminary analyses of the review results, arranged by the six categories included in the review:

* Study
* Objectives
* Taxa and location
* Survey methods
* Parameters and covariates
* Modelling methods

```{r Read-in, echo = FALSE}
reviewResults <- read_rds("Analysis/Outputs/ReviewResults.rds")
```

## Study 
Basic information on the locations and affiliations of authors involved with each paper was collected. Two thirds of papers (36/54) were by authors from U.S. institutions; the next greatest total was 3 papers from Australia.

```{r Study, echo = FALSE}
studyData <- reviewResults[["Study"]]
```

```{r Study authorship countries, echo = FALSE, message = FALSE}
countriesShapefile <- st_read("Analysis/SpatialData/Countries/World_Countries__Generalized_.shp")

## Map of authorship countries by # of papers
studyData %>%
  # Get count of papers by author's first country
  group_by(`First author country`) %>%
  summarise(Count = n()) %>%
  arrange(-Count) %>% 
  mutate(`First author country` = case_when(`First author country` == "United States of America" ~ "United States",
                                            `First author country` != "United States of America" ~ `First author country`)) %>%
  # Merge by country name
  right_join(countriesShapefile, by = c('First author country' = 'COUNTRY')) %>%
  # Plot map
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = log(Count))) +
  labs(title = "Log paper count by country") +
  theme_void() + 
  scale_fill_continuous(name = "Log counts", low="thistle2", high="darkred",na.value="white")

# Table with the same data
studyData %>%
  group_by(`First author country`) %>%
  summarise(`Total papers` = n()) %>%
  arrange(-`Total papers`) %>%
  kable()
```

Affiliated institutions were split into four categories: Academic, Non-governmental organisations, Governmental agencies, and private sector groups. These were fairly varied, with academic groups participating in the vast majority of papers but with substantial representation from governmental and NGO groups.

```{r Affiliation types, echo = FALSE}
# Calc % of each type
studyData %>%
  summarise(NGO = sum(Author_NGO)/n()*100,
            Academic = sum(Author_Academic)/n()*100,
            Government = sum(Author_Government/n()*100),
            Private = sum(Author_Private)/n()*100) %>% 
  # Make legible for table
  pivot_longer(cols = c(NGO, Academic, Government, Private), names_to = 'Affiliation', values_to = "% of papers") %>%
  arrange(-`% of papers`) %>%
  kable(digits = 1)
```

## Objectives
Five categories of objectives and goals were assessed for each paper:

* Assessing trends - interested in the trajector of estimated or derived parameters
* Testing hypotheses - evaluating attributes of a **specific** covariate on parameters
* Spatial mapping - Extrapolating estimates of papameter values beyond surveyed sites
* Future projections - Making predictions of parameters into the future
* Methods - introducing, testing, or demonstrating a new extension to these models

```{r Objectives, echo = FALSE}
objData <- reviewResults[["Objectives"]]
```

Monitoring trends and testing hypotheses were the most popular objectives; making predictions (spatial or temporal) were quite rare.
```{r Objectives table, echo = FALSE}
# Make table 
objData %>% 
  pivot_longer(cols = c(Obj_Methods, Obj_Projections, Obj_Hypothesis, Obj_Spatial, Obj_Trends),
               names_to = "Objective", values_to = "TF") %>%
  # Drop obj_ prefix
  mutate(Objective = str_replace(Objective, "Obj_","")) %>%
  # Tally up and make table
  group_by(Objective) %>%
  summarise(`Percent of papers` = 100*sum(TF)/54) %>%
  arrange(-`Percent of papers`) %>%
  kable(digits = 1)
```

```{r Obj by Authorship group, echo = FALSe}
objByAuth <- full_join(objData, studyData, by = c("Review ID", "Title")) %>%
  # Pivot authorship column and adjust  
  pivot_longer(cols = c(Author_NGO, Author_Academic, Author_Government, Author_Private), 
               names_to = 'Authorship', values_to = "Auth_TF") %>%
  mutate(Authorship = str_replace(Authorship, "Author_","")) %>%
  # Get count of # of papers by authorship for the plot later
  filter(Auth_TF == TRUE) %>%
  group_by(Authorship) %>%
  mutate(nAuth = n()) %>% 
  # Pivot objectives column and adjust
  pivot_longer(cols = c(Obj_Methods, Obj_Projections, Obj_Hypothesis, Obj_Spatial, Obj_Trends),
               names_to = "Objective", values_to = "Obj_TF") %>%
  mutate(Objective = str_replace(Objective, "Obj_","")) %>%
  filter(Obj_TF == TRUE)

objByAuth %>%
  filter(Obj_TF == TRUE & Auth_TF == TRUE) %>%
  group_by(Authorship) %>%
  mutate(TotalAuth = length(unique(Title))) %>%
  group_by(Authorship, Objective, nAuth) %>%
  summarise(Count = n()/nAuth) %>%
  ungroup() %>%
  complete(Authorship, Objective) %>%
  mutate(Count = case_when(is.na(Count) ~ 0.01,
                           !is.na(Count) ~ Count)) %>%

  ggplot(aes(x = Authorship, y = Count, fill = Objective)) +
  geom_col(position = "dodge") +
  ylim(0, 1) + 
  labs(title = "Percent of papers by affiliation pursuing each objective")
```

```{r Covs by obj, echo = FALSE}
covData <- reviewResults[["Covariates"]] %>%
  mutate(`Review ID` = as.numeric(str_replace(`Review ID`, "[:alpha:]", "")))

objData %>%
   pivot_longer(cols = c(Obj_Methods, Obj_Projections, Obj_Hypothesis, Obj_Spatial, Obj_Trends),
               names_to = "Objective", values_to = "Obj_TF") %>%
  mutate(Objective = str_replace(Objective, "Obj_","")) %>%
  filter(Obj_TF == TRUE) %>%
  select(`Review ID`, Objective) %>%
  full_join(covData, by = 'Review ID') %>%

  group_by(`Review ID`, Objective, Parameter) %>%
  summarise(Count = n()) %>%
  filter(Parameter %in% c("Colonisation", "Extinction", "InitialOccupancy", "Detection"),
         Count < 20) %>%

  ggplot(aes(x = Objective, y = Count, colour = Parameter)) +
  geom_boxplot()

```

## Taxa and location
Details on the study organism was collected for each study to assess any trends in the types of species these models are applied to, with a particular focus on those that are vulnerable to conservation threats. The locations where surveys were conducted was also noted.

```{r Taxa, echo = FALSE}
taxaData <- reviewResults[["Taxa"]]
```

As with authorship, the United States dominated the countries where surveys took place.

### Study site maps {.tabset}
#### Country
```{r Site country map, echo = FALSE}
taxaData %>%
  # Divide by unique title/country
  group_by(Country, Title) %>%
  summarise() %>%
  # Get count of papers by site country
  group_by(Country) %>%
  summarise(Count = n()) %>%
  arrange(-Count) %>% 
  mutate(Country = case_when(Country == "United States of America" ~ "United States",
                             Country != "United States of America" ~ Country)) %>%
  # Merge by country name
  right_join(countriesShapefile, by = c('Country' = 'COUNTRY')) %>%
  # Plot map
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = log(Count))) +
  labs(title = "Log paper count by country") +
  theme_void() + 
  scale_fill_continuous(name = "Log counts", 
                        low = "thistle2", 
                        high = "darkred",
                        na.value = "white")

# Table with the same data
taxaData %>%
  # Divide by unique title/country
  group_by(Country, Title) %>%
  summarise() %>%
  # Get count of papers by site country
  group_by(Country) %>%
  summarise(`Total papers` = n()) %>%
  arrange(-`Total papers`) %>% 
  kable()
```

#### Continent
```{r Site continent map, echo = FALSE}
continentShapefile <- st_read("Analysis/SpatialData/Continents/World_Continents.shp")

taxaData %>%
  # Divide by unique title/country
  group_by(Country, Continent, Title) %>%
  summarise() %>%
  # Get count of papers by site continent
  group_by(Continent) %>%
  summarise(Count = n()) %>%
  arrange(-Count) %>% 
  mutate(Continent = case_when(Continent == "Oceania" ~ "Australia",
                               Continent != "Oceania" ~ Continent)) %>%

  # Merge by continent name
  right_join(continentShapefile, by = c('Continent' = 'CONTINENT')) %>%
  # Plot map
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = Count)) +
  labs(title = "Paper count by continent") +
  theme_void() + 
  scale_fill_continuous(name = "Log counts", 
                        low = "thistle2", 
                        high = "darkred",
                        na.value = "white")

# Table with the same data
taxaData %>%
  # Divide by unique title/country
  group_by(Country, Title, Continent) %>%
  summarise() %>%
  # Get count of papers by continent
  group_by(Continent) %>%
  summarise(`Total papers` = n()) %>%
  arrange(-`Total papers`) %>% 
  kable()

```

#### Realm
```{r Site realm map, echo = FALSE}
#ecoregionsShapefile <- st_read("Analysis/SpatialData/Ecoregions/Terrestrial_Ecoregions.shp") %>%
#  st_make_valid() 
#
#t <- ecoregionsShapefile %>%
#  group_by(WWF_REALM2) %>%
#  summarise(geometry = st_combine(geometry))
#
#taxaData %>%
#  # Divide by unique title/country
#  group_by(Country, Realm, Title) %>%
#  summarise() %>%
#  # Get count of papers by site continent
#  group_by(Realm) %>%
#  summarise(Count = n()) %>%
#  arrange(-Count) %>% 
#  
#  # Merge by realm
#  right_join(ecoregionsShapefile, by = c('Realm' = 'WWF_REALM2')) %>%
#  # Plot map
#  ggplot() +
#  geom_sf(aes(geometry = geometry, fill = Count)) +
#  labs(title = "Paper count by Udvardy realm") +
#  theme_void() + 
#  scale_fill_continuous(name = "Log counts", 
#                        low = "thistle2",
#                        high = "darkred",
#                        na.value="white")
#
# Table with the same data
taxaData %>%
  # Divide by unique title/country
  group_by(Country, Title, Realm) %>%
  summarise() %>%
  # Get count of papers by site realm
  group_by(Realm) %>%
  summarise(`Total papers` = n()) %>%
  arrange(-`Total papers`) %>% 
  kable()
```

## Survey methods

```{r Survey, echo = FALSE}

```

## Parameters and covariates

```{r Covariates, echo = FALSE}

```

## Modelling decisions

```{r Modelling, echo = FALSE}

```