### 2023-11-27 Manuscript/index.qmd

---
title: Dynamic occupancy models review
date: today
authors:
  - name: Saoirse Kelleher
    affiliation: University of Melbourne
    email: saoirse.kelleher@student.unimelb.edu.au
    orcid: 0000-0002-4614-5753
    corresponding: true
  - name: Natalie Briscoe
    affiliation: University of Melbourne
  - name: Damaris Zurell
    affiliation: University of Potsdam
  - name: Marc Kery
    affiliation: Vogelwarte
  - name: Gurutzeta Guillera-Arroita
    affiliation: Consejo Superior de Investigaciones Científicas (CSIC)
  - name: Jane Elith
    affiliation: University of Melbourne

abstract: >
  Dynamic occupancy models allow researchers to estimate species occupancy across space and time while accounting for imperfect detection.

bibliography: references.bib
---

## Introduction

## Methods

## Results

## Discussion

## Conclusions