### 2023-11-27 Randomisation/Randomise_Papers.qmd
completed randomisation document

---
title: Randomise papers
subtitle: Describes random selection of papers for inclusion in the review
author: Saoirse Kelleher
date: today
format: 
  html:
    code-fold: true
    theme: minty
    toc: true
bibliography: references.bib
---

## Overview

As this review evolved, two samples of target papers were acquired from two distinct sampling protocols.

-   The **First Sample** was acquired in October 2021 from Google Scholar. The first 100 hits for each of four search terms ('dynamic occupancy model,' 'occupancy dynamics model,' 'multi-season occupancy model,' and 'stochastic patch occupancy model', plus allowances for pluralisation) were downloaded. These were stratified by their query and year-window (2000-2005, 2006-2010, 2010-2015, and 2016-2021). A random sample of 25% of the strata *or* 5 papers, whichever was larger, was flagged for inclusion in the review, to be replaced from the same strata if inclusion criteria were not met.

-   After reviewing this initial sample, the need for a **Second Sample** was identified as many of the papers included in the first did not represent our model of interest. This second, more narrowly targeted sample was drawn from papers citing @mackenzie2003, the original description of the model class which all subsequent implementations would be expected to cite. Papers were acquired via Web of Science in September 2023. The same 4 year-window strata were used here, with 15 papers randomly drawn from each strata. A fifth strata was added for 2022-2023, to capture the most recent papers on these models.

For transparency and reproducibility, the randomisation code is preserved in this document.

## First sample

The full list of papers acquired from the aforementioned queries is read into R, and each is assigned to a year-window.

```{r load packages}
#| message: false
library("knitr")
library("tidyverse") 
library("readxl")

set.seed(5378)

# From the excel review spreadsheet, read in only the identifying characteristics 
first_idSheet <- readxl::read_xlsx("Spreadsheets/Stage1_Spreadsheet.xlsx", 
                                     sheet = "Metadata", skip = 1) %>%
  # Drop rows prior to start of data
  slice(4:nrow(.)) %>%
  # Only need google search query and the ID reference
  select(Review_ID, Query = `Search terms`)

# Repeat with the study metadata
first_studyDataSheet <- readxl::read_xlsx("Spreadsheets/Stage1_Spreadsheet.xlsx", 
                                    sheet = "Study data", skip = 1) %>%
  # Cut out non-data rows
  slice(4:nrow(.)) %>%
  # Only need year data and ID reference
  select(Review_ID, Title, Year)

# Merge paper queries and metadata
first_paperList <- full_join(first_idSheet, first_studyDataSheet, 
                             by = "Review_ID") %>%
  # Exclude some non-query papers and add year periods to the rest
  # Drop empty rows, excluded, and incidental papers
  filter(!is.na(Query) & Query != "INCIDENTAL" & Year != "EXCLUDE") %>%
  # Break into time periods 
  mutate(YearPeriod = case_when(Year %in% as.character(2000:2005) ~ "2000:2005",
                                Year %in% as.character(2006:2010) ~ "2006:2010",
                                Year %in% as.character(2011:2015) ~ "2011:2015",
                                Year %in% as.character(2016:2021) ~ "2016:2021"))

head(first_paperList) %>%
  kable()
```

A random number is then assigned to *all* papers. Papers are then ranked by number within their year-window/query strata. Papers in the top 5/25% (whichever was larger) were flagged to be read, and the rest were kept available for replacement of papers which may fail to meet inclusion criteria.

```{r Assign random numbers}
# Randomly rank papers, then rank within the strata
first_rankedPapers <- first_paperList %>%
  # Add random ranks for all papers
  mutate(randRank = sample(x = 1:nrow(.), size = nrow(.))) %>%
  # Re-rank within strata
  group_by(Query, YearPeriod) %>%
  arrange(Query, YearPeriod, randRank) %>%
  mutate(GroupRank = row_number()) %>%
# Identify which papers qualify to be read
  group_by(Query, YearPeriod) %>%
  mutate(PercentRank =  GroupRank/max(GroupRank)) %>%
  mutate(ToRead = case_when(
    # If there are fewer than 5, read them all
    max(GroupRank) <= 5 ~ "YES",
    # If there are 20 or fewer, read the top 5
    max(GroupRank) <= 20 & GroupRank <= 5 ~ "YES",
    max(GroupRank) <= 20 & GroupRank >= 5 ~ "NO",
    # If there are 21 or more, read the top 25%
    max(GroupRank) >= 21 & PercentRank <= 0.25  ~ "YES",
    max(GroupRank) >= 21 & PercentRank > 0.25 ~ "NO")) %>%
  ungroup() %>%
  unite(col = Group, c(Query, YearPeriod)) %>%
  # Get Review_ID, GroupRank, and ToRead
  select(Review_ID, Title, Group, OverallRank = randRank, GroupRank, ToRead)

head(first_rankedPapers) %>%
  kable()
```

Incidental papers which were not ranked are returned to the dataframe, and the output is saved.

```{r re-add unranked papers}
first_unrankedPapers <- first_paperList %>% 
  # Drop empty rows, excluded, and incidental papers
  filter(is.na(Query) | Query == "INCIDENTAL" | Year == "EXCLUDE") %>%
  mutate(ToRead = "NO", Group = "Unranked", GroupRank = "NA", OverallRank = "NA") %>%
  select(Review_ID, Title, Group, OverallRank, GroupRank, ToRead)

# Combine ranked and unranked papers
firstSample <- rbind(first_rankedPapers, first_unrankedPapers) %>%
  mutate(Review_ID = as.numeric(Review_ID)) %>%
  arrange(Review_ID) %>%
  mutate(Sample = "First")

# Write to disk
write_csv(firstSample, file = "Randomisation/FirstSampleRanks.csv")
```

## Second sample

The list of papers pulled from Web of Science is loaded in - 1184 papers are represented here. Year-windows are assigned here as in the **First Sample.**

```{r Load second sample}
#| warning: false
# Load query and extract relevant columns
second_Query <- readxl::read_xls("Queries/MacKenzie WOS/MacKenzie03.xls") %>%
  select(Review_ID, Title = `Article Title`, Year = `Publication Year`) %>%
  # Assign year period
  mutate(YearPeriod = case_when(Year %in% 2000:2005 ~ "2000:2005",
                                Year %in% 2006:2010 ~ "2006:2010",
                                Year %in% 2011:2015 ~ "2011:2015",
                                Year %in% 2016:2021 ~ "2016:2021",
                                Year %in% 2022:2023 ~ "2022:2023"))

head(second_Query) %>%
  kable()
```

Papers are assigned a random number and ranked within their groups, and top-15 papers are flagged to be read. The output is saved for future reference.

```{r randomise second sample}
secondSample <- second_Query %>%
  # Add query flag to match with the first sample
  mutate(Query = "MacKenzie Cited") %>%
  # Add random ranks for all papers
  mutate(randRank = sample(x = 1:nrow(.), size = nrow(.))) %>%
  # Re-rank within strata
  group_by(Query, YearPeriod) %>%
  arrange(Query, YearPeriod, randRank) %>%
  mutate(GroupRank = row_number()) %>% 
  mutate(ToRead = case_when(
    GroupRank <= 15 ~ "YES",
    GroupRank >15 ~ "NO")) %>%
  unite(col = Group, c(Query, YearPeriod)) %>%
  mutate(Sample = "Second") %>%
  # Get Review_ID, GroupRank, and ToRead
  select(Sample, Review_ID, Title, Group, OverallRank = randRank, GroupRank, ToRead)

write_csv(secondSample, "Randomisation/SecondSampleRanks.csv")

head(secondSample) %>%
  kable()
```

## Sample summary

A basic representation of the two sample's coverage is presented here. The data presented is from after randomisation, **not** after reviewing, and thus does not represent the final coverage of the review as some papers may be excluded.

### First sample categories

```{r plot first sample categories}
firstSample %>%
  separate(Group, c("Query", "YearWindow"), sep = "_") %>%

  ggplot() +
  geom_bar(aes(x = YearWindow, fill = ToRead)) +
  scale_fill_manual(values = c("gray70", "dodgerblue2")) +
  facet_wrap(vars(Query), nrow = 2) +
  labs(title = "Coverage of the Primary Sample",
       y = "Number of Papers",
       x = "Year Window",
       fill = "Included")
```

### Second sample categories

```{r plot second sample categories}
secondSample %>%
  separate(Group, c("Query", "YearWindow"), sep = "_") %>%

  ggplot() +
  geom_bar(aes(x = YearWindow, fill = ToRead)) +
  scale_fill_manual(values = c("gray70", "firebrick3")) +
  labs(title = "Coverage of the Secondary Sample",
       y = "Number of Papers",
       x = "Year Window",
       fill = "Included")
```

### Combined sample

Adding the second sample improves coverage across the review period, particularly in the first two year-windows where fewer papers were included in the first sample.

```{r combined sample}
rbind(firstSample, secondSample) %>%
  separate(Group, c("Query", "YearWindow"), sep = "_") %>%
  filter(ToRead == "YES") %>%

  ggplot() +
  geom_bar(aes(x = YearWindow, fill = Sample)) +
  scale_fill_manual(values = c("dodgerblue2", "firebrick3")) +
  labs(title = "Coverage of combined sample",
       y = "Number of papers",
       x = "Year-window")
```

## References