### 2023-11-27 Analysis/Archive/Code/Archive/2_PaperFigures.Rmd

---
title: "DOM review: Paper figures"
output: html_document
---

```{r Packages, include = FALSE}
library(tidyverse)
library(readxl)
library(knitr)
library(sf)
library(ggalluvial)
```

# Read-in data
```{r read in data, echo = FALSE}
reviewResults <- read_rds("Analysis/Outputs/ReviewResults.rds")
```

# Map of study sites
```{r map of study sites}
# load world map
continents <- st_read("Analysis/SpatialData/Continents/World_Continents.shp")

sitePoints <- reviewResults$Taxa %>%
  select(Title, `Center point`) %>%
  separate(`Center point`, into = c('Latitude', 'Longitude'), sep = ", ") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = st_crs(continents))

ggplot() +
  geom_sf(data = continents, aes(fill = CONTINENT), show.legend = FALSE) +
  geom_sf(data = sitePoints, colour = "black", shape = 4, size = 4) +
  theme_void()

```

```{r alluvial plot}
reviewResults$Modelling %>% 
  filter(Bayesian %in% c("Y", "N")) %>%
  mutate(Bayesian = case_when(Bayesian == "Y" ~ "Bayesian",
                              .default = "Frequentist")) %>%
  group_by(`Prefiltering category`, `Selection category`, Bayesian) %>%
  summarise(n = n()) %>%

  ggplot(aes(y = n, axis1 = `Prefiltering category`, axis2 = `Selection category`)) +
  geom_alluvium(aes(fill = `Selection category`)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 7) +
  facet_grid(rows = vars(Bayesian), scales = "free_y") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 20),
        strip.text = element_text(size = 20))

```

```{r}

vec <-  c("test_a20", "test2_n15", "testing3_p3")

str_replace(string = vec, pattern = "(?<=\\_)[:alpha:]{1}", replacement = "s")

```

```{r stacked bar covs}
simpleCovs <- reviewResults$Covariates %>%
  select(Title, Parameter, CovCategory, EnvCategory, Scale, CovObsType) %>%
  left_join(select(reviewResults$Study, Title, Year), by = "Title")

simpleCovs %>%
  mutate(Parameter = case_when(Parameter == "Persistence" ~ "Extinction",
                               .default = Parameter)) %>%
  filter(Parameter %in% c("InitialOccupancy", "Colonisation", "Extinction", "Detection")) %>%
  filter(CovCategory != "NONE") %>%
  filter(!CovCategory %in% c('UNCLEAR', 'UNSPECIFIED')) %>%
  group_by(Parameter, CovCategory) %>% 
  summarise(avgcount = n()/47) %>%

  ggplot() +
  geom_col(aes(x = Parameter, y = avgcount, fill = CovCategory)) +
  scale_fill_discrete("Covariate Category",
                      limits = c("E", "CONN", "INT", "SEAS", "SPEC", "SURV", "OTHER"),
                     labels = c("Environmental", "Connectivity", "Interactions", "Season effect",
                               "Species effect", "Survey details", "Other")) +
  scale_x_discrete(limits = c("InitialOccupancy", "Colonisation", "Extinction", "Detection")) +
  labs(title = "Average number of covariates per parameter", y = "Average number of covariates", x = "Parameter")

# Total number of covariates by year
simpleCovs %>%
  filter(Parameter %in% c("InitialOccupancy", "Colonisation", "Extinction", "Detection")) %>%
  filter(CovCategory != "NONE") %>%
  filter(!is.na(CovCategory)) %>%
  filter(!CovCategory %in% c('UNCLEAR', 'UNSPECIFIED')) %>%
  filter(!is.na(Year)) %>% 
  group_by(Year) %>%
  mutate(npapers = length(unique(Title))) %>%
  ungroup() %>%
  group_by(CovCategory, Year, npapers) %>%
  summarise(count = n()) %>%
  mutate(avgCount = count/npapers) %>%

  ggplot() + 
  geom_col(aes(x = Year, y = avgCount, fill = CovCategory)) +
  scale_fill_discrete("Covariate Category",
                      limits = c("E", "CONN", "INT", "SEAS", "SPEC", "SURV", "OTHER"),
                    labels = c("Environmental", "Connectivity", "Interactions", "Season effect",
                               "Species effect", "Survey details", "Other")) +
  labs(title = "Average total number of covariates by year", 
       y = "Average number of covariates")
```

```{r covs by objective}
reviewResults$Objectives %>%
  select(Title, Obj_Methods, Obj_Projections, Obj_Spatial, Obj_Hypothesis) %>%
  pivot_longer(cols = c(Obj_Methods, Obj_Spatial, Obj_Hypothesis, Obj_Projections),
               names_to = "Objective", values_to = "ObjIncluded") %>%
  filter(ObjIncluded == TRUE) %>%
  mutate(Objective = case_when(Objective == "Obj_Methods" ~ "Methods",
                               Objective == "Obj_Spatial" ~ "Spatial Prediction",
                               Objective == "Obj_Projections" ~ "Future Prediction",
                               Objective == "Obj_Hypothesis" ~ "Hypothesis Testing")) %>%

  right_join(simpleCovs, by = "Title") %>%
  filter(Parameter %in% c("InitialOccupancy", "Colonisation", "Extinction", "Detection")) %>%
  filter(CovCategory != "NONE") %>%
  filter(!is.na(CovCategory)) %>%
  filter(!CovCategory %in% c('UNCLEAR', 'UNSPECIFIED')) %>%
  group_by(Objective) %>%
  mutate(npapers = length(unique(Title))) %>%
  ungroup() %>%
  group_by(CovCategory, Objective, npapers) %>%
  summarise(count = n()) %>%
  mutate(avgCount = count/npapers) %>%  
  filter(!is.na(Objective)) %>%

  ggplot() + 
  geom_col(aes(x = Objective, y = avgCount, fill = CovCategory)) +
  scale_fill_discrete("Covariate Category",
                      limits = c("E", "CONN", "INT", "SEAS", "SPEC", "SURV", "OTHER"),
                    labels = c("Environmental", "Connectivity", "Interactions", "Season effect",
                               "Species effect", "Survey details", "Other")) +
  labs(title = "Average total number of covariates by objective", 
       y = "Average number of covariates")

```

```{r stat tables}
reviewResults$Study %>%
  group_by(`First author country`) %>%
  summarise(n = n())

reviewResults$Survey %>%
  group_by(`Capture method`) %>%
  summarise(n = n())

reviewResults$Covariates %>%
  filter(Parameter %in% c("InitialOccupancy", "Colonisation", "Extinction", "Detection")) %>%
  filter(CovCategory != "NONE") %>%
  filter(!is.na(CovCategory)) %>%
  filter(!CovCategory %in% c('UNCLEAR', 'UNSPECIFIED')) %>%
  group_by(CovCategory, EnvCategory, Title) %>%
  summarise() %>%
  group_by(CovCategory, EnvCategory) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(percent = n/47)

```