### 2023-11-27 Queries/QueryComparison.Rmd

---
title: "Web of Science and Google Scholar Comparison"
date: "September 14 2023"
output: html_document
---

```{r Setup, include = FALSE}
library(tidyverse)
library(readxl)
papers <- read_excel("Queries/Query Comparison.xlsx")
```

Find overlap
```{r, Find overlaps}
overlaps <- papers %>%
  mutate(Overlap = Query) %>%
  group_by(Title, Query, Year) %>%
  mutate(Overlap = if_else(n()>1, TRUE, FALSE)) %>%
  mutate(Platform = case_when(Overlap == TRUE ~ "Both",
                              .default = Platform)) %>%
  group_by(Title, Query, Year, Platform) %>%
  summarise()
```

```{r overlaps}
ggplot(overlaps) +
  geom_bar(aes(x = Query, fill = Platform))

ggplot(overlaps) +
  geom_bar(aes(x = Year, fill = Platform))

ggplot(overlaps) +
  geom_bar(aes(x = Year, fill = Platform)) +
  facet_wrap(vars(Query))

```