### 2024-03-14 Analysis/ReviewSummary/ReviewSummary.qmd
started outlining sections and data availability

---
title: "DOM Review Summary"
author: "Saoirse Kelleher"
date: "March 13 2024"
date-modifed: today
abstract: "A summary of results from the DOM Review, for use in interpreting outputs and writing the review manuscript. "
abstract-title: "Summary"
format: 
  html:
    theme: minty
    toc: true
---

```{r prepare inputs}
#| include: false

library(tidyverse)
library(readxl)

# Study Details
review_StudyDetails <- read_xlsx("Spreadsheets/Stage3_Spreadsheet.xlsx",
                                 sheet = "Study") |>
  select(ID = "Review ID", Sample, Status, Title, Journal, Year, 
         FirstAuthor = "First author country", 
         AuthorCategories = "Authorship groups")

# Objectives and Inference
review_Objectives <- read_xlsx("Spreadsheets/Stage3_Spreadsheet.xlsx",
                               sheet = "Objectives and Inference") |>
  select(ID = "Review ID", Status, Title, 
         Obj_Methods = "Methods?",
         Obj_Future = "Future predictions?",
         Obj_Spatial = "Spatial predictions?", 
         Obj_Hypothesis = "Hypothesis testing?")

# Species and Location
review_SpeciesLocation <- read_xlsx("Spreadsheets/Stage3_Spreadsheet.xlsx",
                                    sheet = "Species and Location") |>
  select(ID = "Review ID", Status, Title, 
         Taxa_Category = "Taxa keywords", Taxa_Number = "Number",
         IUCN_Status = "IUCN Status", Threat_Category = "Threat categories",
         Country, Continent, Realm, Size_km = "Study Size")

# Data Collection
review_DataCollection <- read_xlsx("Spreadsheets/Stage3_Spreadsheet.xlsx",
                                   sheet = "Data collection") |>
  select(ID = "Review ID", Status, Title,
         Duration = "Study duration", Detection_Method = "Capture method",
         Primary_Occasions = "Number of primary occasions",
         Secondary_Occasions = "Number of secondary occasions",
         Secondary_Type = "Type of secondary occasions",
         Site_Number = "Number of sites", Site_Type = "Type of site",
         Sampling_Method = "Sampling method")

# Covariates
review_Covariates <- read_xlsx("Spreadsheets/Stage3_Spreadsheet.xlsx",
                               sheet = "Covariates") |>
  select(ID = "Review ID", Status, Title, 
         Parameter, Covariate_Count = "Total covariates",
         Covariates = "Covariate List")

# Modelling
review_Modelling <- read_xlsx("Spreadsheets/Stage3_Spreadsheet.xlsx",
                              sheet = "Modelling") |>
  select(ID = "Review ID", Status, Title,
         Bayesian, Prefiltering = "Prefiltering category",
         Selection = "Selection category", Evaluation = "Evaluation method",
         Multispecies = "Multi-species", Connectivity)

```

# Study details

Basic metadata was collected for each paper in the review, with available data including:

| Field             | Description                                      | Possible values                      |
|--------------|------------------------------------|----------------------|
| Journal           | Journal which this article was published in      | *Any*                                |
| Year              | Year of publication                              | *Any*                                |
| Author country    | Nation of the first author's primary affiliation | *ISO 3166 country names*             |
| Author categories | Categories of affiliations for *all* authors     | *Academic, Government, NGO, Private* |

## Publication journals

## Authorship countries

## Authorship categories

## Authorship category trends

## Review coverage

# Objectives and inference

The stated objectives for each article were assessed and divided into four categories - each paper could fall into multiple (or zero) objective categories.

| Objective           | Description                                                                       |
|---------------------|---------------------------------------------------|
| Methods             | Did authors seek to demonstrate or test specific methods for occupancy modelling? |
| Future Predictions  | Did authors make predictions beyond the study period?                             |
| Spatial Predictions | Did authors make predictions at locations where surveys were not conducted?       |
| Hypothesis testing  | Did authors seek to test specific hypotheses of species occupancy dynamics?       |

## Frequency of objectives

## Stated objectives by authorship categories

## Stated objective by year

# Species and location

Data on focal species and the location of study areas was collected, available fields include:

| Field           | Description                                                                                | Possible values                                                                   |
|----------------|----------------------------|----------------------------|
| Taxa category   | Category of study taxa; could fit into multiple categories                                 | *Bird, Mammal, Invertebrate, Herptile, Fish, Threatened, Invasive, Other*         |
| Taxa number     | Number of individual species within the taxa for which models were fit                     | *Numeric*                                                                         |
| IUCN status     | IUCN status for the study species; NA if multi-species or if the study species is invasive | *LC, NT, EN, CR, DD, NA*                                                          |
| Threat category | Category of threats to the species; multiple per species permitted                         | *Habitat loss, Climate change, Exploitation, Invasives, Disease, Disaster, Other* |
| Country         | Country/countries where the study took place                                               | *ISO 3166 country names*                                                          |
| Continent       | Continent(s) where the study took place                                                    | *North America, South America, Europe, Asia, Africa, Oceania*                     |
|                 |                                                                                            |                                                                                   |
|                 |                                                                                            |                                                                                   |
|                 |                                                                                            |                                                                                   |
|                 |                                                                                            |                                                                                   |
|                 |                                                                                            |                                                                                   |

# Data collection

##

# Covariates

# Modelling