### 2024-04-12 Manuscript/domReview.qmd

---
title: "Twenty years of dynamic occupancy modelling --- what’s been done, and where to next?"
date: today
authors:
  - name: Saoirse Kelleher
    affiliation: University of Melbourne
    email: saoirse.kelleher@student.unimelb.edu.au
    orcid: 0000-0002-4614-5753
    corresponding: true
  - name: Natalie Briscoe
    affiliation: University of Melbourne
  - name: Gurutzeta Guillera-Arroita
    affiliation: Pyrenean Institute of Ecology, Spanish National Research Council
  - name: Jane Elith
    affiliation: University of Melbourne

bibliography: references.bib
embed-resources: true
---

# Introduction

Since its introduction two decades ago by @mackenzie2003, the dynamic multi-season occupancy model has become an important tool for analysis of species occupancy throughout space and time. Dynamic occupancy models (DOMs) represent a happy medium in the ecological modelling landscape. In explicitly estimating the processes which guide occupancy (i.e., colonisation and extinction) this model class offers more advanced insights than simpler static models of species distribution which have noted deficits in their ability to describe non-equilibrium states **\[cite\]**. The capacity to account for imperfect detection provides a strong additional benefit given the near-ubiquity of this source of bias in ecological field data. Even with these advantages, data input requirements for DOMs remain achievable for many research programs, particularly compared to other more complex process-explicit models.

Potential use cases of this model class are wide-ranging. Simple estimates of occupancy through time have value in many monitoring programs, particularly for threatened or invasive species where distributions are likely to exist in a non-equilibrium state **\[cite?\]**. These models can also be used in hypothesis testing of specific relationships, with examples including the effect of fire on extinction probability and the influence of species interactions on occupancy **\[cite,cite\]**. A less frequently pursued yet promising application of DOMs lies in their ability to make predictions, either to unsurveyed areas or into the future [@briscoe2021; @kéry2013a].

This review aims to provide an overview of the first two decades of the dynamic occupancy models, including its development and how authors have implemented it in their investigations of ecological questions. We present key guidelines for authors seeking to build these models in their work, and outline the key priorities and opportunities for future research incorporating this flexible and important model class.

### The dynamic occupancy model

Dynamic occupancy models jointly describe two processes: the ecological process of site occupancy describing the presence or absence of a species at a given site in a given point in time; and the observational process of detection describing whether a species is observed given that a site is occupied. The basic model form as described by @mackenzie2003 is depicted in @fig-modelform. Under this parameterisation sites exist in either occupied or unoccupied states which they may transition between through time. Occupancy state in the first time-step is given by the initial occupancy parameter, and in ensuing time-steps occupancy is described as a Markovian process governed by probabilities of colonisation and extinction. Given that a site is occupied, the probability of observing the species during a survey is given by the detection parameter. Under this formulation, it is assumed that false-positive detections do not occur.

![The original parameterisation of the dynamic occupancy model. The ecological process of occupancy is guided by the initial occupancy, colonisation, and extinction parameters; the observational process is guided only by the detection parameter.](Figures/ModelForm.png){#fig-modelform width="1000"}

To disentangle the ecological and observation processes and estimate the detection parameter, the DOM relies on a hierarchical sampling design shown in @fig-surveys. Under this design, observations at a site occur during distinct, time-bound seasons in which sites are considered 'closed' to changes in occupancy. Within seasons multiple observations are conducted, which permit estimation of the probability of detection conditional on occupancy. These repeat observations may be conducted by a variety of methods. Most frequently surveys are simply repeated on separate occasions, although alternative methods exist. These include surveys conducted at multiple locations within a site **\[cite\]**, use of multiple observers during a survey **\[cite\]**, or by recording time elapsed before a detection is recorded **\[cite\]**.

![The sampling design of the dynamic occupancy model.](Figures/SurveyDesign.png){#fig-surveys fig-align="center" width="589"}

Dynamic occupancy models require a few assumptions with important implications for users.

1.  **False positive detections do not occur.** While this assumption can be safely met in some studies, it is not guaranteed when working with more cryptic species. Several authors have commented on the perils of not accounting for false positives where they occur (@miller2015, @mcclintock2010). Furthermore, @berigan2019 notes that in some cases genuine identifications of a species may be considered false positives if they do not represent the right kind of occupancy - in this case, transient individuals where the authors were interested in breeding occupany. Extensions to the dynamic occupancy model which account for false positive error are available (@royle2006, @miller2011, @miller2013) and should be carefully considered where this assumption cannot be met.

2.  **Sites are closed to changes in occupancy between seasons.** This assumption, best known as the 'closure assumption,' has also been subject to considerable discussion on the bias which is introduced when it is violated (@rota2009, @otto2013). Closure is dependent not only on the behaviour of the species, but also the definition of occupancy used by researchers — short seasons may represent dynamics more representative of species 'use,' and @valente2017 discusses the difficultly in distinguishing temporary emigration and local extinction. Model extensions to relax the closure assumption have been developed, including @kendall2013 's approach using staggered arrival and depature periods between sites. More pertinent for most studies, however, is a careful consideration of ecologically and observationally relevant seasons which will allow the estimation of an appropriate 'type' of occupancy, whether that be shorter or longer term.

3.  **No modelled heterogeneity in ecological or observational parameters.** Relative to the former two assumptions, considerably less research has been conducted on the impact of unmodelled heterogeneity on parameter estimates. Particularly at larger spatial scales, environmental factors are near-guaranteed to influence species occupancy and observation.

# Methods

This review is focused on the dynamic, multi-season, implementation of the occupancy model of @mackenzie2003. To qualify for inclusion articles must fit a model which fulfils these criteria: i) multiple sites which can exist in at least two states, including an occupied and unoccupied states; ii) multiple survey periods, between which sites may change state conditional on the prior survey period's occupancy state and at least one transition probability; iii) contains at least one detection parameter; iv) uses real (i.e., not simulated) field ecological data.

Papers were solicited for inclusion in two stages. In a preliminary analysis, four search strings ("dynamic occupancy model," "occupancy dynamics model," "incidence function model," and "multi-season occupancy model" were fed into Google Scholar **\[cite\]**. Up to 100 articles were downloaded for each string in order of appearance. Articles were divided into four temporal strata: 2000-2005, 2006-2010, 2011-2015, and 2016-2021. For each combination of search string and year strata, either a random 25% or 5 articles (whichever was greater) were processed. Articles which did not meet inclusion criteria were replaced with another from the same search string and strata where possible.

After completion of the preliminary sample authors recognised significant uncertainty in how the model of interest was described, particularly in their first decade of use, and noted that the incidence-function model is a distinct model class **\[cite\].** To increase sample size, a complete list of articles which have cited @mackenzie2003 was downloaded from Web of Science. Articles in this list were split into the same year strata, plus an additional 2022-2023 category. 15 articles were drawn from each strata and again processed with replacement from their strata.

Details on methods for each article were recorded in detail, including the full suite of covariates considered for inclusion in modelling. The full spreadsheet of reviewed articles is available in **Appendix I**.

# Results

A total of **\[number\]** articles were included in the review, **\[number\]** from the preliminary sample and **\[number\]** from the secondary sample. Based on the acceptance rate from the secondary sample, an estimated **\[hit number\]** of the **\[total number\]** articles citing @mackenzie2003 would have met inclusion criteria; i.e., fit a dynamic occupancy model.

![](Figures/CoveragePlot.jpeg){#fig-coverage}

## Study details

Articles were published with first-authors from a total of **\[number\]** countries, with authors from the United States producing **\[zz\]**% of articles. **\[x\]**% of articles had at least one author based at an academic institution, **\[y\]**% had authors at government institutions, and \[**z\]**% had authors from non-profit organisations.

Data for DOMs were collected from study areas on all continents and in **\[xx\]** countries, with the United States again representing a majority of the total articles published (@fig-StudyDetails a). These study areas ranged in size from less than 1 to over 1 million square kilometres

![](Figures/ReviewBox.png){#fig-StudyDetails}

## Data collection

## Modelling

![](Figures/CovParamPlot.jpeg){#fig-covariates}

## Selection

![](Figures/SelectionFigure.jpeg){#fig-selection}

### Complexity

![](Figures/ModelComplexity.png){#fig-complexity}

# Discussion

**\<Covariate inclusion figure\>**

We provide guidelines for fitting dynamic occupancy models.

![](Figures/Guidelines.png){#fig-guidelines}

#