### 2024-04-12 Manuscript/domReview.qmd
Complete sections
completed manuscript sections rough text

---
title: "Twenty years of dynamic occupancy modelling --- what’s been done, and where to next?"
date: today
authors:
  - name: Saoirse Kelleher
    affiliation: University of Melbourne
    email: saoirse.kelleher@student.unimelb.edu.au
    orcid: 0000-0002-4614-5753
    corresponding: true
  - name: Natalie Briscoe
    affiliation: University of Melbourne
  - name: Gurutzeta Guillera-Arroita
    affiliation: Pyrenean Institute of Ecology, Spanish National Research Council
  - name: Jane Elith
    affiliation: University of Melbourne

bibliography: references.bib
embed-resources: true
---

# Introduction

Since its introduction two decades ago by @mackenzie2003, the dynamic multi-season occupancy model has become an important tool for analysis of species occupancy throughout space and time. Dynamic occupancy models (DOMs) represent a happy medium in the ecological modelling landscape. In explicitly estimating the processes which guide occupancy (i.e., colonisation and extinction) this model class offers more advanced insights than simpler static models of species distribution which have noted deficits in their ability to describe non-equilibrium states **\[cite\]**. The capacity to account for imperfect detection provides a strong additional benefit given the near-ubiquity of this source of bias in ecological field data. Even with these advantages, data input requirements for DOMs remain achievable for many research programs, particularly compared to other more complex process-explicit models.

Potential use cases of this model class are wide-ranging. Simple estimates of occupancy through time have value in many monitoring programs, particularly for threatened or invasive species where distributions are likely to exist in a non-equilibrium state **\[cite?\]**. These models can also be used in hypothesis testing of specific relationships, with examples including the effect of fire on extinction probability and the influence of species interactions on occupancy **\[cite,cite\]**. A less frequently pursued yet promising application of DOMs lies in their ability to make predictions, either to unsurveyed areas or into the future [@briscoe2021; @kéry2013a].

This review aims to provide an overview of the first two decades of the dynamic occupancy models, including its development and how authors have implemented it in their investigations of ecological questions. We present key guidelines for authors seeking to build these models in their work, and outline the key priorities and opportunities for future research incorporating this flexible and important model class.

### The dynamic occupancy model

Dynamic occupancy models jointly describe two processes: the ecological process of site occupancy describing the presence or absence of a species at a given site in a given point in time; and the observational process of detection describing whether a species is observed given that a site is occupied. The basic model form as described by @mackenzie2003 is depicted in @fig-modelform. Under this parameterisation sites exist in either occupied or unoccupied states which they may transition between through time. Occupancy state in the first time-step is given by the initial occupancy parameter, and in ensuing time-steps occupancy is described as a Markovian process governed by probabilities of colonisation and extinction. Given that a site is occupied, the probability of observing the species during a survey is given by the detection parameter. Under this formulation, it is assumed that false-positive detections do not occur.

![The original parameterisation of the dynamic occupancy model. The ecological process of occupancy is guided by the initial occupancy, colonisation, and extinction parameters; the observational process is guided only by the detection parameter.](Figures/ModelForm.png){#fig-modelform width="1000"}

To disentangle the ecological and observation processes and estimate the detection parameter, the DOM relies on a hierarchical sampling design shown in @fig-surveys. Under this design, observations at a site occur during distinct, time-bound seasons in which sites are considered 'closed' to changes in occupancy. Within seasons multiple observations are conducted, which permit estimation of the probability of detection conditional on occupancy. These repeat observations may be conducted by a variety of methods. Most frequently surveys are simply repeated on separate occasions, although alternative methods exist. These include surveys conducted at multiple locations within a site **\[cite\]**, use of multiple observers during a survey **\[cite\]**, or by recording time elapsed before a detection is recorded **\[cite\]**.

![The sampling design of the dynamic occupancy model.](Figures/SurveyDesign.png){#fig-surveys fig-align="center" width="589"}

Dynamic occupancy models require a few assumptions with important implications for users.

1.  **False positive detections do not occur.** While this assumption can be safely met in some studies, it is not guaranteed when working with more cryptic species. Several authors have commented on the perils of not accounting for false positives where they occur (@miller2015, @mcclintock2010). Furthermore, @berigan2019 notes that in some cases genuine identifications of a species may be considered false positives if they do not represent the right kind of occupancy - in this case, transient individuals where the authors were interested in breeding occupany. Extensions to the dynamic occupancy model which account for false positive error are available (@royle2006, @miller2011, @miller2013) and should be carefully considered where this assumption cannot be met.

2.  **Sites are closed to changes in occupancy between seasons.** This assumption, best known as the 'closure assumption,' has also been subject to considerable discussion on the bias which is introduced when it is violated (@rota2009, @otto2013). Closure is dependent not only on the behaviour of the species, but also the definition of occupancy used by researchers — short seasons may represent dynamics more representative of species 'use,' and @valente2017 discusses the difficultly in distinguishing temporary emigration and local extinction. Model extensions to relax the closure assumption have been developed, including @kendall2013 's approach using staggered arrival and depature periods between sites. More pertinent for most studies, however, is a careful consideration of ecologically and observationally relevant seasons which will allow the estimation of an appropriate 'type' of occupancy, whether that be shorter or longer term.

3.  **No modelled heterogeneity in ecological or observational parameters.** Relative to the former two assumptions, considerably less research has been conducted on the impact of unmodelled heterogeneity on parameter estimates. Particularly at larger spatial scales, environmental factors are near-guaranteed to influence species occupancy and observation.

# Methods

## Article solicitation

This review is focused on the dynamic, multi-season, implementation of the occupancy model of @mackenzie2003. To qualify for inclusion articles must fit a model which fulfils these criteria: i) multiple sites which can exist in at least two states, including an occupied and unoccupied states; ii) multiple survey periods, between which sites may change state conditional on the prior survey period's occupancy state and at least one transition probability; iii) contains at least one detection parameter; iv) uses real (i.e., not simulated) field ecological data.

Papers were solicited for inclusion in two stages. In a preliminary analysis, four search strings ("dynamic occupancy model," "occupancy dynamics model," "incidence function model," and "multi-season occupancy model" were fed into Google Scholar **\[cite\]**. Up to 100 articles were downloaded for each string in order of appearance. Articles were divided into four temporal strata: 2000-2005, 2006-2010, 2011-2015, and 2016-2021. For each combination of search string and year strata, either a random 25% or 5 articles (whichever was greater) were processed. Articles which did not meet inclusion criteria were replaced with another from the same search string and strata where possible.

After completion of the preliminary sample authors recognised significant uncertainty in how the model of interest was described, particularly in their first decade of use, and noted that the incidence-function model is a distinct model class **\[cite\].** To increase sample size, a complete list of articles which have cited @mackenzie2003 was downloaded from Web of Science. Articles in this list were split into the same year strata, plus an additional 2022-2023 category. 15 articles were drawn from each strata and again processed with replacement from their strata.

Details on methods for each article were recorded in detail, including the full suite of covariates considered for inclusion in modelling. The full spreadsheet of reviewed articles is available in **Appendix I**.

## Analysis

For each included article we collected data collected on authorship, stated research objectives, study taxa and field survey locations, data collection methods, covariates, and model selection and evaluation. Particular emphasis was placed on the covariates considered by authors, with no distinction between covariates which were or were not included in final models. As we were principally interested in the modelling process, our analysis did not directly explore final models or the results generated by those models. A spreadsheet of reviewed articles and collected data is provided in Appendix I.

# Results

A total of **\[number\]** articles were included in the review, **\[number\]** from the preliminary sample and **\[number\]** from the secondary sample. Based on the acceptance rate from the secondary sample, an estimated **\[hit number\]** of the **\[total number\]** articles citing @mackenzie2003 would have met inclusion criteria. Total published articles have increased considerably over time.

![Estimated coverage of articles included in this review](Figures/CoveragePlot.jpeg){#fig-coverage}

Articles were published with first-authors from a total of **\[number\]** countries, with authors from the United States producing **\[zz\]**% of articles. **\[x\]**% of articles had at least one author based at an academic institution, **\[y\]**% had authors at government institutions, and \[**z\]**% had authors from non-profit organisations.

## Study locations, focal species, and data collection

Data for DOMs were collected from study areas on all continents and in **\[xx\]** countries, with the United States again representing a majority of the total articles published. These study areas ranged in size from less than 1 to over 1 million square kilometres (@fig-StudyDetails a).

Focal species were largely vertebrate taxa, although 6 articles fit models to invertebrates. A single non-animal taxa was represented in the sample, from an article fitting a DOM to a rare species of lichen @belinchón2017. [aa%] of articles modelled threatened taxa, and \[bb%\] modelled species invasive to the study area. The majority of articles (50, X%) fit single-species DOMs to one taxa. Of the remainder, 26 (X%) fit independent single-species models to more than one taxa. 8 articles fit (X%) multi-species implementations with explicit interactions between species (@fig-StudyDetails b).

Data used to fit dynamic occupancy models included human survey data such as transects or point counts (X% of articles), live-trapping data (X%), camera traps (X%) or bioacoustic monitors (Y%), and citizen science data (Z%); some articles incorporate multiple detection methods to build their detection datasets. Study duration, defined as the time elapsed between the first and last surveys conducted, ranged from 1 month to 40 years with the median study running for X years. The number of seasons ranged from 2 to Y with a median of Z; the number of observations per season ranged from X to Y with a median of Z. The number of sites surveyed per article ranged from X% to Y%, with the median study including Z sites (@fig-StudyDetails c).

![](Figures/ReviewBox.png){#fig-StudyDetails}

## Covariates

A summary of the covariates classes considered in the review sample is presented in @tbl-covariates. The most frequent

| Covariate type | Proportion of articles considering | Proportion including on Initial Occupancy | Proportion dynamic | Proportion directly observed | Frequency of interaction terms | Frequency of polynomial terms |
|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
|                |                                    |                                           |                    |                              |                                |                               |
|                |                                    |                                           |                    |                              |                                |                               |
|                |                                    |                                           |                    |                              |                                |                               |
|                |                                    |                                           |                    |                              |                                |                               |
|                |                                    |                                           |                    |                              |                                |                               |
|                |                                    |                                           |                    |                              |                                |                               |
|                |                                    |                                           |                    |                              |                                |                               |
|                |                                    |                                           |                    |                              |                                |                               |
|                |                                    |                                           |                    |                              |                                |                               |
|                |                                    |                                           |                    |                              |                                |                               |

: Covariates considered in dynamic occupancy models {#tbl-covariates}

![](Figures/CovParamPlot.jpeg){#fig-covariates}

### Complexity

![](Figures/ModelComplexity.png){#fig-complexity}

### Selection and evaluation

Bayesian models were far less likely to implement a model selection protocol compared to frequentist models: 30% of the former fit only a single model defined *a priori,* compared to 10% of the latter (@fig-selection). In frequentist models, stepwise selection protocols were the most popular choice. The vast majority of frequentist models (XX%) used AIC or AICc to identify the best model.

![](Figures/SelectionFigure.jpeg){#fig-selection}

Objective evaluation of model fit was rare throughout the sample, regardless of selection protocol. XX% of articles calculated non-relative metrics, either via in-sample validation, bootstrapped goodness-of-fit tests, or out-of-sample validation.

# Discussion

### Usage of dynamic occupancy models

Our review sample displays considerable diversity in implementations of dynamic occupancy models with widespread representation from mammals, birds, and herptiles. This assemblage contrasts with the related stochastic patch occupancy models (SPOMs, @gutiérrez-arellano), where the majority of models are fit to invertebrates, plants, and fungi. This may reflect SPOMs lack of consideration for imperfect detection and assumption of discrete habitat patches, limitations which are unlikely to be fulfilled by most vertebrate taxa.

DOMs also display great variation in scale, from local studies with a handful of sites to continental-scale studies spanning decades.

Several types of detection data have been successfully incorporated DOMs. This flexibility may be one of the more notable aspects of dynamic occupancy models; while the heirarchical data inputs may be more complex than those required by other models of species distributions, authors have been able to successfully manipulate datasets from many sources to meet requirements.

### Practices in the modelling process

Model selection approaches diverge between Bayesian and frequentist implementations of the model. While model selection will naturally look different between these two categories, it is unclear what effect a lack of model selection may have on the outputs produced by Bayesian models and it is likely that Bayesian model selection approaches should be further explored @hooten2015.

Notably, the popular sequential approach to model selection (where one sub-model is assessed at a time) may not provide optimal results as shown in @morin2020. Considerably more research is required on this topic to clarify how authors should best approach the model building process.

### Key considerations for model users

Implementation of dynamic occupancy models requires considerable thought to ensure that model outputs correspond with author expectations and ecological realities. Based upon our review of the existing literature we discuss key questions authors should carefully engage with during the modelling process.

1.  **Are seasons ecologically meaningful, and how do they affect the definition of occupancy?** As previously discussed, the duration of the season has important implications for both the reliability and proper interpretation of model results. Seasons which are not ecologically meaningful may run afoul of the closure assumption resulting in associated biases. Also relevant is how the definition of 'occupancy' may shift with season length and life-history of a species — for many species, occupancy within a year and within a week may be considerably different concepts, with shorter intervals more closely relating to the idea of 'use.' With the increasing proliferation of continuous detection methods such as camera traps and acoustic monitors where the definition of 'season' is to some degree arbitrary, it is increasingly important to thoughtfully consider how this element of the DOM is defined.

2.  **Are key drivers of occupancy and detection incorporated?** While it is effectively impossible to *fully* describe heterogeneity across a landscape, authors should make their best efforts to capture the drivers of occupancy and detection most likely to influence their study species. This is true even when another covariate is of principal interest to investigators — failure to include other major determinants of occupancy is likely to bias parameter estimates. One place where this is of particular importance is in determining drivers of detection, which is not only subject to observational conditions but also ecological determinants which may cause the species to spend more time at a location and thus be more frequently available for detection.

3.  **Ensure covariates reflect ecological complexity.** Many ecological relationships are non-linear and may require polynomials terms to be accurately represented. This is particularly true at large spatial scales where a larger portion of a species ecological niche is shown in surveyed data. Authors should question whether a linear term is sufficient to capture hypothetical relationships in their study systems and consider quadratic terms where appropriate.

4.   **Consider model evaluation.** Model evaluation remains rare in ecological research relative to other fields embracing complex modelling, likely due to a reluctance to 'throw away' data for use in testing. This concept should be challenged, particularly for larger scale long-term projects where reserving a fraction of data for testing is unlikely to substantially reduce precision of parameter estimates. For studies where this is genuinely not possible, reporting results of goodness-of-fit tests may still provide valuable information to readers.

### Priorities for future research

Work on dynamic occupancy modelling over the past 20 years has developed this model class into an important tool for describing species distributions and ecological relationships. Historically, methods-development research for dynamic occupancy models has focused on questions of model structure — namely ideas around the closure assumption and imperfect detection. This emphasis likely reflects the DOM's roots in the mark-recapture literature, where these topics have been discussed for several decades. Considerably less emphasis has been placed on the relation of model parameters to environmental factors via covariates and the model selection processes used to do so. This stands in stark contrast to related correlative species distribution models, where these topics have been the subject of considerable research over the years. There are opportunities for the DOM literature to learn from species distribution models, and more targeted research on the influence of the model selection process is necessary.

The use of DOMs in prediction has remained fairly limited despite significant potential for this use-case. Papers in our sample which have taken advantage of this potential included @mcgowan2020 who used DOMs to make predictions of black rail distributions under various management scenarios, and @saunders2020, who validated predictions from ecological niche models using continent-scale DOMs. Further implementations of these models may benefit from the aforementioned necessary research on model selection to ensure robust outputs.

A final opportunity for future research relates to autonomous detection methods, such as camera traps and bio-acoustic monitors — these technologies have permitted collection of enormous quantities of data particularly well-suited to dynamic occupancy models, but have also raised additional questions. The definition of seasons is of increased importance for these data types, where division of data into 'seasons' and 'observations' is arbitrary. Further research should explore the implications of this phenomenon and additional modelling opportunities made possible by this novel data type.

The past two decades have shown that dynamic occupancy models are an excellent tool for addressing a wide range of ecological questions relating to species occupancy throughout space and time. The flexibility afforded by model extensions and the considerable advantages of the class over other species distribution models is likely to ensure that these models continue to increase in popularity. Given their strong use on threatened taxa and excellent potential to support environemental management, it is important that research continues to support model users in creating the most reliable models possible.