### 2024-09-23 Randomisation/Randomisation.qmd

---
title: Randomise papers for DOM Review
subtitle: Describes random selection of papers for inclusion in the review
author: Saoirse Kelleher
date: today
format: 
  html:
    code-fold: true
    theme: minty
    toc: true
    embed-resources: true
---

```{r packages}
#| message: false

options(tidyverse.quiet = TRUE)
library("knitr")
library("tidyverse")
library("readxl")
```

## Load Web of Science queries

Two search queries were made on Web of Science on July 26 2024.

1.  ALL=("dynamic\* occupancy model\*") OR TS=(dynamic\* NEAR "occupancy model\*") OR TI=(dynamic\* NEAR "occupancy model\*") OR AB=(dynamic\* NEAR "occupancy model\*") OR ALL=("occupancy dynamic\*") OR ALL=("multi\*season occupancy model\*") OR ALL=("multi-season occupancy model\*") OR (ALL=(occupancy) AND ALL=(coloni?ation) AND (ALL=(extinction) OR ALL=(persistence)) AND ALL=(detect\*)) and Ecology or Biodiversity Conservation or Environmental Sciences or Zoology or Ornithology or Evolutionary Biology or Biology or Marine Freshwater Biology or Forestry or Fisheries or Entomology or Plant Sciences or Environmental Studies (Web of Science Categories)
2.  All articles citing MacKenzie et al. 2003

WOS datasheets containing all matching records are loaded in.

```{r load WOS queries}
keyTerms_datasheet <- read_xlsx("Queries/Final Queries/Keywords.xlsx",
                                skip = 10) |>
  select(Title, Year = "Publication Year", Journal = "Source Title", 
         Authors, Citations = "Total Citations", DOI) |>
  mutate(Query = "KeyTerms")

citesMackenzie_datasheet <- read_xlsx("Queries/Final Queries/MacKenzie_A.xlsx",
                                      skip = 10) |>
  rbind(read_xlsx("Queries/Final Queries/MacKenzie_B.xlsx",
                  skip = 10)) |>
  select(Title, Year = "Publication Year", Journal = "Source Title", 
         Authors, Citations = "Total Citations", DOI) |>
  mutate(Query = "CitesMackenzie")
```

The datasheets are then combined, and duplicate records are removed. All records prior to 2004 are also removed. Each article is given a unique ID.

```{r combine articles}
combined_datasheet <- rbind(keyTerms_datasheet,
                            citesMackenzie_datasheet) |>
  filter(Year %in% 2004:2023) |>
  mutate(QueryValue = TRUE) |>
  pivot_wider(names_from = Query, values_from = QueryValue,
              values_fill = FALSE) |>
  mutate(Review_ID = 1:n())
```

## Add year strata and randomise

```{r randomise articles}
withr::with_seed(14512, {
  randomised_datasheet <- combined_datasheet |>
    mutate(Strata = case_when(Year %in% 2004:2007 ~ "2004-2007",
                              Year %in% 2008:2011 ~ "2008-2011",
                              Year %in% 2012:2015 ~ "2012-2015",
                              Year %in% 2016:2019 ~ "2016-2019",
                              Year %in% 2020:2023 ~ "2020-2023")) |>
    mutate(StrataRank = sample(1:n(), n()),
           .by = Strata) |>
    arrange(Strata, StrataRank) |>
    select(Review_ID, StrataRank, Strata, Title, Year, 
           Journal, Authors, Citations, DOI, 
           KeyTerms, CitesMackenzie)
})
```

## Write randomised datasheets

```{r Save sheet}
write_csv(randomised_datasheet, "Randomisation/Final_Randomisation.csv")
```